\chapter*{Abstract}

\addcontentsline{toc}{chapter}{Abstract}
As education becomes increasingly modular and lifelong learning gains prominence, traditional credential verification systems remain limited to authenticating whole degrees or certificates. Current blockchain implementations reduce administrative burden but fail to capture the granular nature of learning achievements, creating opportunities for credential forgery through timeline manipulation and fabrication of learning histories. This research introduces IU-MiCert, a blockchain system that enhances existing credential management through verifiable academic micro-credential tracking with authentic timeline verification. Using advanced cryptographic data structures that improve storage and verification efficiency, IU-MiCert provides detailed learning record management while significantly strengthening protection against credential fraud through comprehensive timeline tracking. The system supplements existing credential verification by representing learning as a continuous, verifiable journey with cryptographically secured chronological records. This detailed academic history makes credential forgery substantially more difficult by creating comprehensive audit trails that expose inconsistencies in fabricated educational progressions. IU-MiCert enables students to build verifiable profiles incrementally, with each course, project, and achievement securely recorded with authentic timestamps within their broader educational context. For educational institutions, the system provides periodic update cycles that efficiently manage detailed learning records while minimizing blockchain storage costs. For business, especially employers, it enables verification of specific skills and authentic learning progression beyond single degree completion, supporting the increasingly important paradigm of skills-based hiring and lifelong learning verification. By enhancing traditional credential systems with detailed academic provenance, IU-MiCert transforms credential verification from basic authentication into comprehensive demonstration of authentic learning pathways, providing greater transparency, trust, and fraud protection for all stakeholders in the educational ecosystem.

\vspace{0.5cm}

\noindent\textbf{Keywords:} micro-credentials, academic provenance, blockchain verification, credential fraud prevention, learning timeline, verkle tree